---
title: "Clean data: finalize"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# This file provides the startup() function.
source("R/_startup.R")

# Load desired packages and report any missing packages that should be installed.
startup(auto_install = FALSE, verbose = FALSE)

# Load any additional R files in the R/ directory.
ck37r::load_all_code("R", verbose = TRUE)
```

Previous file: 2-clean-impute.Rmd

## Load data

```{r load_data}
# Created in 2-clean-impute.Rmd
# Objects included: data, vars
# renv also includes a load() method, so we specify base:: here.
base::load("data/clean-impute.RData")
```

## Factors to indicators

```{r factors_indicators}
result = ck37r::factors_to_indicators(data[vars$predictors], verbose = TRUE)

names(result)

# Temporarily remove all predictors from the dataframe.
data[vars$predictors] = NULL

# Now add the new data back on.
data = cbind(data, result$data)

# Remove the original factor predictors from the list of used predictors, add
# add the new indicator predictors.
(vars$predictors = c(setdiff(vars$predictors, result$factor_vars), unlist(result$factor_names)))

# Confirm that our predictor vector is updated correctly.
if (!all(vars$predictors %in% names(data))) {
  missing_vars = setdiff(vars$predictors, names(data))
  stop("Missing new indicators that were added: ", paste(missing_vars, collapse = ", "))
}
  

rm(result)
```

## Remove collinear predictors

This is not essential, but nice for the linear regression estimators.

This needs to be after imputation, because it currently cannot handle missingness.

```{r remove_collinear}
# Remove linearly correlated columns from the covariate file
# NOTE: assumes that there are no factor variables.
linear_combos = caret::findLinearCombos(data[, vars$predictors])

if (length(linear_combos$remove) > 0L) {

  if (conf$verbose) {
    cat("Removing", length(linear_combos$remove), "predictors due to collinearity.\n")
    cat("Vars:", paste0(vars$predictors[linear_combos$remove], collapse = ", "), "\n")
  }
  
  # Make sure we don't switch to a vector if only 1 column remains.
  data = data[, !colnames(data) %in% vars$predictors[linear_combos$remove],
              drop = FALSE]
  
  vars$predictors = setdiff(vars$predictors, vars$predictors[linear_combos$remove])
  
  if (conf$verbose) {
    cat("Updated predictor count:", length(vars$predictors), "\n")
  }
} else {
  cat("No linear duplication found.\n")
}

rm(linear_combos)
```

## Confirm predictor matrix invertability

This is not essential, but nice for the linear regression estimators.

```{r confirm_invertability}
# Compute covariance matrix.
# NOTE: this requires that no factors be included.
cov_mat = stats::cov(data[vars$predictors])
  
# Compute QR decomposition of covariance matrix.
qr_cov = base::qr(cov_mat)

# These need to be equal for the covariance matrix to be full rank.
if (ncol(cov_mat) != qr_cov$rank) {
  cat("Warning: matrix of predictors is not full rank.\n")
  cat("Predictor columns:", ncol(cov_mat), "QR rank:", qr_cov$rank, "\n")
} else {
  cat("Predictor matrix is full rank.\n")
}
rm(cov_mat, qr_cov)
```

## Save finalized dataset

```{r save_finalized}
save(data, vars,
     file = "data/clean-finalize-imputed.RData")
```

Next file: 4-eda.Rmd